\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-11-17 rebuttal.tex

}
\opening{Dear chair and members of the reading committee}
\todo{fix todos in document}

I would like to extend my sincere gratitude for time and effort you have taken
to review the manuscript of my PhD dissertation.
Additionally, I would like to thank you for the insightful, constructive and
detailed comments you all have provided.
The following reply contains a point-by-point response to your review comments.
Attached, you will also find an updated draft of the manuscript and a version
with indicated changes for your approval.

\reply{Andrea K\"ubler}
Major comments:
\begin{enumerate}
	\item
\end{enumerate}
Minor comments:
\begin{enumerate}
	\item
\end{enumerate}

\reply{Fabien Lotte}
Major comments:
\begin{enumerate}
	\item Part of the confusion between active and passive BCI might be due
	to the poor choice of phrasing ``paradigms that require little active participation from the user, like redirect-
	ing attention or initiating imagined actions, are classified as passive BCI.''.
	This has been rewritten to ``in contrast to paradigms that require the user's active participation, for
	instance through redirecting attention or initiating imagined actions,
	paradigms free of this requirement are classified as passive bci,''
	for clarity.
	Neurofeedback has been reclassified as active BCI in the text and in
	figure 1.3.
	\item \todo{Why XDAWN+RG no shrinkage (performed worse), fair
	comparison of sample rate}
	\item \todo{report actual accuracies performance stbf in table}
	\item In the results section of chapter 4, statistical comparisons were
	only performed between methods for which we evaluated the performance
	ourselves (HODA, PARAFACDA and BTTDA).
	Since the MOABB benchmarking framework~\cite{Aristimunha2023} allows for
	fair comparison of performance metrics reported across different works,
	the analyses for other comparison methods were not re-run
	by us as reported, but performances were
	taken from the MOABB benchmark database~\cite{Chevallier2024}.
	However, this source provided only the aggregated performance metrics
	and not the metrics per subject, session and cross-validation fold,
	hence we could not perform relevant statistical comparison with these
	comparison models.
	\item \todo{include table with absolute performances for transfer
	conditions}
\end{enumerate}
Minor comments:
\begin{enumerate}
	\item According to the annotations, typographical errors have been
	corrected and imprecise or inaccurate statements have been adjusted.
	The recommended references have been discussed and, where necessary,
	statements were adapted to accommodate this.
\end{enumerate}

### 2024-11-18 rebuttal.tex

}
\opening{Dear chair and members of the reading committee}
\todo{fix todos in document}

I would like to extend my sincere gratitude for time and effort you have taken
to review the manuscript of my PhD dissertation.
Additionally, I would like to thank you for the insightful, constructive and
detailed comments you all have provided.
The following reply contains a point-by-point response to your review comments.
Attached, you will also find an updated draft of the manuscript and a version
with indicated changes for your approval.

\reply{Andrea K\"ubler}
Major comments:
\begin{enumerate}
	\item
\end{enumerate}
Minor comments:
\begin{enumerate}
	\item
\end{enumerate}

\reply{Fabien Lotte}
Major comments:
\begin{enumerate}
	\item Part of the confusion between active and passive BCI might be due
	to the poor choice of phrasing ``paradigms that require little active participation from the user, like redirect-
	ing attention or initiating imagined actions, are classified as passive BCI.''.
	This has been rewritten to ``in contrast to paradigms that require the user's active participation, for
	instance through redirecting attention or initiating imagined actions,
	paradigms free of this requirement are classified as passive bci,''
	for clarity.
	Neurofeedback has been reclassified as active BCI in the text and in
	figure 1.3.
	\item \todo{Why XDAWN+RG no shrinkage (performed worse), fair
	comparison of sample rate}
	\item \todo{report actual accuracies performance stbf in table}
	\item In the results section of chapter 4, statistical comparisons were
	only performed between methods for which we evaluated the performance
	ourselves (HODA, PARAFACDA and BTTDA).
	Since the MOABB benchmarking framework~\cite{Aristimunha2023} allows for
	fair comparison of performance metrics reported across different works,
	the analyses for other comparison methods were not re-run
	by us as reported, but performances were
	taken from the MOABB benchmark database~\cite{Chevallier2024}.
	However, this source provided only the aggregated performance metrics
	and not the metrics per subject, session and cross-validation fold,
	hence we could not perform relevant statistical comparison with these
	comparison models.
	\item \todo{include table with absolute performances for transfer
	conditions}
\end{enumerate}
Minor comments:
\begin{enumerate}
	\item According to the annotations, typographical errors have been
	corrected in text and figures, and imprecise or inaccurate statements have been adjusted.
	The recommended references have been discussed and, where necessary,
	statements were updated accordingly.
  Figure 1.5 was corrected and matched with the correct caption.
\end{enumerate}

### 2024-11-18 chapters/stbf_struct.tex
Rework fig 1.3 and 1.4
Update figure 5

  \begin{figure}
		\includegraphics[width=.66\linewidth]{figures/stbf_struct/accuracy.eps}
    \caption[Clasifier accuracy in function of available training data.]{%
      Accuracy of the different classifiers for all 21 subjects relative to the
			number of blocks available for training. One block consists of 135
			epochs and corresponds to 27 seconds of stimulation. Accuracies
			are shown for the evaluation settings averaging over 1, 2, and
			3 trials of testing stimuli.
			\cref{fig:stbf-struct/accuracy-ap} contains results for all numbers of trials.
			While \textsc{stbf-emp} is unstable when little training data
			are available,
			regularization of the covariance matrix (\textsc{stbf-shrunk} and
			\textsc{stbf-struct}) drastically improves performance.}
		\label{fig:accuracy}
	\end{figure}
  \todo{the legend does not correspond to the text (method names are
  different)}

  \fullpagefig{%
		\includegraphics[width=\textwidth]{figures/stbf_struct/accuracy_app.eps}
  }{%
    \caption[Accuracy averaging over different numbers of trials.]{%
      Accuracy of the different classifiers for all 21 subjects relative to the
		  number of blocks available for training. One block consists of 135
		  epochs and corresponds to 27 seconds of stimulation. Accuracies
		  are shown for the evaluation settings averaging over different
	    numbers of trials, ranging from 1 to 15.
    }
    \label{fig:stbf-struct/accuracy-ap}
  }

@@

	\begin{figure}
		\includegraphics[width=.66\linewidth]{figures/stbf_struct/training_time.eps}
    \caption[Classifier training time.]{Median fold training time for different classifiers at different spatial
			resolution levels evaluated over all training folds for all subjects.
			Error bars represent standard deviation. The training time of \textsc{stbf-shrunk} increases more
			steeply with resolution compared to \textsc{stbf-struct}.
			All \textsc{stbf} classifiers can be trained significantly faster than
			\textsc{xdawn+rg}.}
		\label{fig:stbf-struct/training-time}
	\end{figure}
  \todo{the legend does not correspond to the text (method names are
  different)}

	\cref{fig:stbf-struct/training-time} shows that the training time of
	\textsc{stbf-struct} increases less steeply than that of \textsc{stbf-shrunk}
	and \textsc{stbf-emp}. The training time of all three \ac{stbf}-based classifiers is
  much lower\todo{qualitative} than that of \textsc{xdawn+rg}, which appears nearly constant when using
	4, 8, 16, or 32 channels.

>>>>

  \begin{figure}
    %\includegraphics[width=.66\linewidth]{figures/stbf_struct/accuracy.eps}
    \sffamily
    \sansmath
    \input{figures/stbf_struct/accuracy.pgf}
    \caption[Clasifier accuracy in function of available training data.]{%
      Accuracy of the different classifiers for all 21 subjects relative to the
			number of blocks available for training. One block consists of 135
			epochs and corresponds to 27 seconds of stimulation. Accuracies
			are shown for the evaluation settings averaging over 1, 2, and
			3 trials of testing stimuli.
			\cref{fig:stbf-struct/accuracy-ap} contains results for all numbers of trials.
			While \textsc{stbf-emp} is unstable when little training data
			are available,
			regularization of the covariance matrix (\textsc{stbf-shrunk} and
			\textsc{stbf-struct}) drastically improves performance.}
		\label{fig:accuracy}
	\end{figure}

  \fullpagefig{%
    \sffamily
    \sansmath
    \input{figures/stbf_struct/accuracy_all.pgf}
  }{%
    \caption[Accuracy averaging over different numbers of trials.]{%
      Accuracy of the different classifiers for all 21 subjects relative to the
		  number of blocks available for training. One block consists of 135
		  epochs and corresponds to 27 seconds of stimulation. Accuracies
		  are shown for the evaluation settings averaging over different
	    numbers of trials, ranging from 1 to 15.
    }
    \label{fig:stbf-struct/accuracy-ap}
  }

@@

	\begin{figure}
		%\includegraphics[width=.66\linewidth]{figures/stbf_struct/training_time.eps}
    \sffamily
    \sansmath
    \input{figures/stbf_struct/training_time.pgf}
    \caption[Classifier training time.]{Training time boxplots of different
    classifiers and individual results for all training folds for all subjects
    using 32 channels. Using structured covariance estimation greatly reduces
    training time.}
		\label{fig:stbf-struct/training-time}
	\end{figure}
  \todo{the legend does not correspond to the text (method names are
  different)}

	\cref{fig:stbf-struct/training-time} shows that the training time of
	\textsc{stbf-struct} has a lower median training time than
  \textsc{stbf-shrunk}, \textsc{stbf-emp} and \textsc{xdawn+rg} when training
  on all available data and using 32 channels\todo{qualitative}.
